---
title: "Peer Assigment 1"
author: "Sean Treadway"
date: "June 14, 2015"
output: html_document
---

Report for Peer Assignment #1 in Reproduceable Research

The data to be analysed is pedometer reading generated by an anonymouns individual.  The data was gathered over 2 months, and the actual "step" are binned in to 5-minute interval.  There are a total of 17568 5-minute interval records.  Each record consisted of a "step", "data", and "interval" entry.

The data was provided as part of the assignment.  Data file is "activity.csv", and is in a simple CSV format with 1 header.

Load necessary packages
```{r,echo=TRUE}
library("markdown", lib.loc="/Library/Frameworks/R.framework/Versions/3.2/Resources/library")
library("lattice", lib.loc="/Library/Frameworks/R.framework/Versions/3.2/Resources/library")
library("knitr", lib.loc="/Library/Frameworks/R.framework/Versions/3.2/Resources/library")
library("htmltools", lib.loc="/Library/Frameworks/R.framework/Versions/3.2/Resources/library")
library("rmarkdown", lib.loc="/Library/Frameworks/R.framework/Versions/3.2/Resources/library")
```

Set working directory that contains data and 'figure' folder.
```{r, echo=TRUE}
setwd("~/Documents/DS_track/Repo/peer1/peer1")
```


Here the data is read in to R.
```{r, echo=TRUE}
#Set number of record/lines.  Given in instructions, checked with 'wc' at command line
records=17568
#Set the colume classes
classes<-c("numeric","Date","numeric")
#Set column names.  Comes from header
cnames=c("steps","date","interval")
#Read in data
alldata<-read.delim("./activity.csv",header=TRUE,sep=",", nrows=records,colClasses=classes,strip.white=TRUE)
```

The first part of the assigment was to determine the mean, and median daily steps, and to plot a histogram of daily steps.
NAs in the step entries were ignored for this part of the analysis.
```{r, echo=TRUE}
#Create a logical vector for selecting records with out step NAs
goodrows<-!is.na(alldata$steps)
#Create total steps per day vector
daily<- as.vector(tapply(alldata[goodrows,]$steps,alldata[goodrows,]$date,sum))
#Determine the mean and median daily steps, ignoring NAs
dailymean<-mean(daily)
dailymedian<-median(daily)
dailymean
dailymedian
```
The mean total steps was 10766.189, and the median 10765.

Just below is the histogram for total daily steps.  NA were omitted.
```{r, echo=TRUE}
#Create histogram showing daily step count frequency
hist(daily,breaks=10,main="Histogram of Daily Step Count",xlab="Steps/Day",ylab="Number of Days")
```

The greatest number of step was typically taken between 835 and 840 AM, with about 206 steps on average for the entire 2 month period.  NAs were omitted.
```{r, echo=TRUE}
#Determine the mean for each interval.  This will be used to fill in the step NAs.  
#This totally ignores day/week/month differences that might exist, and any trends in activity. 
#Ignoring these simplifies the analysis for the sake of the exercise.
intmeans24<-tapply(alldata$steps,alldata$interval,mean,na.rm=TRUE)
#Determine what interval sees the highest mean number of steps
max(intmeans24)
which.max(intmeans24)
```

Here is a plot of average steps for each interval.  
```{r, echo=TRUE}
#Plot the mean number of steps on an interval basis
plot(intmeans24,type='l',xlab="Interval (5-minute increment)",ylab="Mean Steps",main="Interval Step Average Across 2 Months")
```

The data is modified to substitute something meaningful in for the 2304 NA step entries.
I chose average for each interval across the 2 month period. Zero (0) was assumed in the case of NA for intervals across the entire time.
```{r, echo=TRUE}
sub<-function(record){
        
        if(is.na(record[1])) {
                record[1]<-intmeans24[as.character(record[2])]
        }
        return(record[1])
}
moddata<-alldata
moddata$steps<-apply(subset(alldata, select=c(steps,interval)),1,sub)

moddaily<- as.vector(tapply(moddata$steps,moddata$date,sum))
#Count of NA values
NAcount<-records-sum(goodrows)
moddailymean<-mean(moddaily)
moddailymean
moddailymedian<-median(moddaily)
moddailymedian
```
The modified daily mean, 10766.189 is basically identical to the unmodified data mean.
The modified daily median is up slightly to 10766.189.

Histogram of modified daily steps is just below.
```{r, echo=TRUE}
hist(moddaily,breaks=10,main="Histogram of Daily Step Count For Modified Data",xlab="Steps/Day",ylab="Number of Days")
```

Modification of the data resulted in basically no change in the data trends.  Day frequency count is up, reflecting a more complete data record.

We would like to explore differences between weekday and weekend step counts.  To do this, a two-level factor is added to the modified data set.  The two levels are "weekday" and "weekend".  The factors were are determined based on the day of the week entry already added to each record.
```{r,echo=TRUE}
#Create a factor for day-of-week section.
moddata$day<-weekdays(moddata$date)

daytype<-function(testday){
        MF<-c('Monday','Tuesday','Wednesday','Thursday','Friday')
        if(testday %in% MF) {
                testday<-"weekday"
        }
        else {
                testday<-"weekend"
        }
        return(testday)
}
```

```{r, echo=TRUE}
moddata$daytype<-as.factor(sapply(moddata$day, daytype))
```

Here, a lattice type plot is create to compare the step count for weekends and weekdays.
```{r,echo=TRUE}

xyplot(steps~interval | daytype ,moddata, layout=c(1,2),aspect='fill', panel=function(x,y){
               panel.xyplot(x,y)
               panel.average(x,y,horizontal = FALSE,col.line = 'red',lwd=5)
       })
```

Using the filled data set, this plot attempts to compare the interval step difference for weekdays to weekends.  The blue circles are all the step counts for the intervals.  The red line plots illustates a moving average for those points.  The averaging seems to incdicate an earlier start on steps at the beginning of the day.  Peak steps happens about the same time for both weekdays and weekends.  Activity through weekend days appears more even, with noticably more activity over the middle of the day as compared to weekdays.